---
title: "Micro Brewery Data Analysis"
author: "Adam Ruthford and Jeff Nguyen"
date: "10/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(class)
library(caret)
library(kableExtra)
```

# Import and clean data
The data for this project was provided by officials from Budweiser brewing company

## Import data
Two data files Beers and Breweries are used in this analysis
```{r ImportData}

# Get raw data
Beers <- read.csv("//DS1513/AllData/Adam/SMU Data Science Courses/DS 6306 Doing Data Science/Unit 8 and 9_ Case Study 1/Beers.csv")
Breweries <- read.csv("//DS1513/AllData/Adam/SMU Data Science Courses/DS 6306 Doing Data Science/Unit 8 and 9_ Case Study 1/Breweries.csv")
```
## Clean Data
Data was examined for missing and duplicate values Cleaning data took the place in several steps

### Duplicate breweries
Two duplicate breweries were found, Summit Brewing Company of St Paul, MN and Sly Fox Brewing Company of Phoenxiville, PA. Beers related to one of the removed breweries were reassigned to the other brewery
```{r DuplicateBreweries}
# Fix duplicate Breweries, Just assign all Beers to one of the duplicates
#Breweries %>% filter(Breweries$Name=="Summit Brewing Company")
Beers[Beers$Brewery_id == 139,]$Brewery_id <- 59

#Breweries %>% filter(Breweries$Name=="Sly Fox Brewing Company")
Beers[Beers$Brewery_id == 164,]$Brewery_id <- 372

# Remove duplicate brewery
Breweries <- Breweries[!(Breweries$Brew_ID == 139 | Breweries$Brew_ID == 164),]
```
### Unique Beers
Several beers were duplicated in the provided data. A unique set of beers was created. For the purposes of this analysis Beer name and brewery id are assumed to be unique
```{r DuplicateBeers, message=FALSE}
UniqueBeers <- Beers %>% group_by(Name,Brewery_id) %>% summarize(minBeer_ID = min(Beer_ID)) %>% select(minBeer_ID)

Beers <- inner_join(Beers, UniqueBeers, by = c("Beer_ID" = "minBeer_ID")) %>%
  select(Name = Name.x, Beer_ID, ABV, IBU, Brewery_id, Style, Ounces)
```
# Analysis Questions
Questions 1,3,4,5,6,7 answered below

## Question 1 : How many breweries are in each state?
```{r Question1}
BreweriesByState <- Breweries %>% select(State) %>% group_by(State) %>% count()
#head(BreweriesByState, 51)
firstBreweries = BreweriesByState[1:17,]
secondBreweries = BreweriesByState[18:34,]
thirdBreweries = BreweriesByState[35:51,]
BreweriesByStateWD <- cbind(firstBreweries,secondBreweries,thirdBreweries)
knitr::kable(BreweriesByStateWD,
             format = "html",
             caption = "Breweries by State",
             col.names = c("State","Brewery Count","State","Brewery Count","State","Brewery Count")) %>% 
  column_spec(1:6, width = "1.1in") %>% 
  row_spec(1:17,align = "center") %>% kable_styling(bootstrap_options = c("hover"))

```
```{r OtherDataByStateCity, include=FALSE}
BreweriesByStateDesc <- BreweriesByState %>% arrange(desc(n))
#BreweriesByCity
BreweriesByCity <- Breweries %>% select(State,City) %>% group_by(State,City) %>% count()
BreweriesByCityDesc <- BreweriesByCity %>% arrange(desc(n))

```
## Question 2 : Data from Beers and Breweries was joined together
```{r JoinBeerBrewery, include=FALSE}
#### AQ 2 ####
BeerAndBrewery <- Breweries %>% inner_join(Beers, by = c("Brew_ID" = "Brewery_id"))
```
## Question 3 : Addressing missing data values
ABV content and IBU rating data is missing for many beers. Mean ABV and IBU data for beers computed by style will be substituted for the missing data

### Compute Mean ABV by Style
```{r Question3ABVStyle}
# Get tables of Mean ABV and IBU by Style. Use those values to replace NA data
ABVByStyle <- Beers %>% 
  dplyr::group_by(Style) %>% 
  dplyr::summarize(ABVStyle = mean(ABV, na.rm = TRUE)) %>% 
  select(Style, ABVStyle) 
```
### Compute Mean IBU by Style
```{r Question3IBUStyle}
IBUByStyle <- Beers %>% 
  dplyr::group_by(Style) %>% 
  dplyr::summarize(IBUStyle = as.integer(round(mean(IBU, na.rm = TRUE), 0))) %>% 
  select(Style, IBUStyle) 
```
### Substitute mean IBU and ABV style data for missing data
```{r Question3}
BeerAndBreweryImproved <- BeerAndBrewery %>% 
  inner_join(ABVByStyle, by = "Style") %>% 
  inner_join(IBUByStyle, by = "Style") %>% 
  mutate(ABV = coalesce(ABV, ABVStyle)) %>%
  mutate(IBU = coalesce(IBU, IBUStyle))
```
## Question 4 : Plot medain alcohol content and IBU rating by state

### Summarize and enhance data for plot
```{r Question4DataManipulation}
BeerAndBreweryImprovedByState <- BeerAndBreweryImproved %>% 
  dplyr::group_by(State) %>% 
  dplyr::summarize(StateABVMean = mean(ABV, na.rm = TRUE), StateABVMedian = median(ABV, na.rm = TRUE), 
                   StateIBUMean = mean(IBU, na.rm = TRUE), StateIBUMedian = median(IBU, na.rm = TRUE),
                   StateBeerCount = n()) %>% 
  select(State, StateABVMean, StateABVMedian, StateIBUMean, StateIBUMedian, StateBeerCount) 

BeerAndBreweryImprovedByState$StateABVMeanPercent <- BeerAndBreweryImprovedByState$StateABVMean * 100.0
BeerAndBreweryImprovedByState$StateABVMedianPercent <- BeerAndBreweryImprovedByState$StateABVMedian * 100.0
```
### Plot the data
```{r Question4Plot, fig.height = 10, fig.width=8}
scaleFactor = 0.1

ggplot(data = BeerAndBreweryImprovedByState, aes(x=State,  width=.4)) +
  geom_col(aes(y=StateABVMedianPercent), fill="blue") +
  geom_col(aes(y=StateIBUMedian * scaleFactor), fill="red", position = position_nudge(x = -.4)) +
  scale_y_continuous(name="Median percent ABV by State", breaks = c(0,1,2,3,4,5,6),
                     sec.axis=sec_axis(~./scaleFactor, name="Median IBU by State", breaks = c(0,10,20,30,40,50,60))) +
  theme(
    axis.title.x.top=element_text(color="red"),
    axis.text.x.top=element_text(color="red"),
    axis.title.x.bottom=element_text(color="blue"),
    axis.text.x.bottom=element_text(color="blue")
  ) +
  coord_flip() +
  labs(x = element_blank()) +
  scale_x_discrete(limits = rev(levels(BeerAndBreweryImprovedByState$State)))
```

## Question 5 : States with the highest ABV and IBU
Unsure of exactly what is being asked. Do we need the state with the highest median ABV and IBU respectively or the states with the beer that has the highest ABV and IBU. The question was worked both ways

### Highest median ABV
```{r Question5MaxMedianABV}
BeerAndBreweryImprovedByState[which.max(BeerAndBreweryImprovedByState$StateABVMedian),] %>% select(State, StateABVMedianPercent)
```

### Highest median IBU
```{r Question5MaxMedianIBU}
BeerAndBreweryImprovedByState[which.max(BeerAndBreweryImprovedByState$StateIBUMedian),] %>% select(State, StateIBUMedian)
```

### Highest ABV for a single beer
```{r Queswtion5MaxBeerABV}
BeerAndBreweryImproved[which.max(BeerAndBreweryImproved$ABV),] %>% select(State, Name.x, Name.y, ABV)
```

### Highest IBU for a single beer
```{r Queswtion5MaxBeerIBU}
BeerAndBreweryImproved[which.max(BeerAndBreweryImproved$IBU),] %>% select(State, Name.x, Name.y, IBU)
```

## Question 6 : Summary and comments on distribution of ABV and IBU variables
Most states have a median ABV for beers brewed of between five and six percent. Two states fall bellow that range UT and NJ, however NJ has only eight different flavors of beer being brewed in that state. Five states have a median ABV for beers brewed above six percent KY, DC, WV, NM, MI, however, DC and WV have less than ten different flavors of beer brewed in them.

## Question 7 : Discuss the relationship between IBU and ABV
The graph below shows the relationship between IBU and ABV, in general as IBU rating increase so does ABV. The only part of the graph that shows a flat line is with very weak beers, those having an alcohol content less than 5 percent. There is a sort of plateau with beers of 7 percent ABV not changing in bitterness. Higher alcohol content beers, those over 8 percent, change bitterness at a slower rate than that of beers from 5 to 6.5 percent ABV.
```{r Question7, fig.height = 6, fig.width=10, message=FALSE,warning=FALSE}
ggplot(data = BeerAndBreweryImproved, mapping = aes(x = ABV * 100, y = IBU)) +
  geom_point(position = "dodge") + geom_smooth(se = FALSE) +
  xlab("Percent ABV") + ylab("IBU Rating") + 
  scale_x_continuous(breaks = c(2,3,4,5,6,7,8,9,10), limits = c(2.5,10)) +
  labs(caption = "ABV values of over 10 percent were left off due to sparseness of the data")
```